---
title: "Assignment_4"
author: "Tejasvini Mavuleti"
date: "3/19/2022"
output:
  word_document: default
  pdf_document: default
---
```{r}
Pharmaceuticals <- read.csv("~/Assignment_4/Pharmaceuticals.csv")
View(Pharmaceuticals)
```

```{r}
getwd()
setwd("C:/Users/mavul/OneDrive/Documents/Assignment_4")
Pharmaceuticaldata<- read.csv("~/Assignment_4/Pharmaceuticals.csv")
str(Pharmaceuticaldata)
```

```{r}
# Calling libraries

library(tidyverse)
library(factoextra)
library(cluster)
library(ggplot2)
library(gridExtra)
```

# a) Use only the numerical variables (1 to 9) to cluster the 21 firms 
```{r}
Pharmaceuticaldata <- na.omit(Pharmaceuticaldata)
Pharmaceuticaldata
row.names(Pharmaceuticaldata)<- Pharmaceuticaldata[,1]
Pharmadata<- Pharmaceuticaldata[, 3:11]
head(Pharmadata)
```

```{r}
# Scaling the Pharmadata using the scale function

Pharma_scale <- scale(Pharmadata)
head(Pharma_scale)
```
```{r}
# Computing K-means clustering and using multiple values of K and examine the difference

km1 <- kmeans(Pharma_scale, centers = 2, nstart = 30)
km2<- kmeans(Pharma_scale, centers = 5, nstart = 30)
km3<- kmeans(Pharma_scale, centers = 6, nstart = 30)
Plot1<-fviz_cluster(km1, data = Pharma_scale)+ggtitle("k=2")
plot2<-fviz_cluster(km2, data = Pharma_scale)+ggtitle("k=5")
plot3<-fviz_cluster(km3, data = Pharma_scale)+ggtitle("k=6")
grid.arrange(Plot1,plot2,plot3, nrow = 3)

distance<- dist(Pharma_scale, method = "euclidean")
fviz_dist(distance)
```

```{r}
# Computing and ploting wss for k = 1 to k = 10 and extracting wss for 2-15 clusters
# The location of a elbow in the plot is considered as an indicator of the number of clusters k =5

set.seed(64060)
wss<- function(k){kmeans(Pharma_scale, k, nstart =10)$tot.withinss}
k.values<- 1:10 
wss_clusters<- map_dbl(k.values, wss)
plot(k.values, wss_clusters, type="b", pch = 16, frame = TRUE, xlab="clusters", ylab="WSS")
```

```{r}
# Final analysis and extracting results using 5 clusters
set.seed(64060)
final_Cluster<- kmeans(Pharma_scale, 5, nstart = 25)
print(final_Cluster)
clusplot(Pharma_scale,final_Cluster$cluster, color = TRUE, labels = 2,lines = 0)
```
# b) Interpret the clusters with respect to the numerical variables used in forming the clusters
Cluster 1 - BAY,IVX,CHTT
Cluster 2 - AGN,PHA
Cluster 3 - ELN,AVE,WPI,MRX
Cluster 4 - BMY,WYE,AHM,ABT,NVS,AZN,LLY
Cluster 5 - JNJ,MRK,GSK,PFE

```{r}
Pharma_Cluster <- Pharmaceuticaldata[,c(12,13,14)]%>% 
mutate(clusters = final_Cluster$cluster)%>% 
arrange(clusters, ascending = TRUE)
Pharma_Cluster
```

# c)Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? 
```{r}
plot1<-ggplot(Pharma_Cluster, mapping = aes(factor(clusters), fill=Median_Recommendation))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
plot2<- ggplot(Pharma_Cluster, mapping = aes(factor(clusters),fill = Location))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
plot3<- ggplot(Pharma_Cluster, mapping = aes(factor(clusters),fill = Exchange))+geom_bar(position = 'dodge')+labs(x ='No of clusters')
grid.arrange(plot1, plot2, plot3)       
```

As per graph, 
Cluster 1- has the highest Beta , leverage and lowest market_cap, ROE, ROA, leverage, rev_growth, net_profit_margin
Cluster 2- has the highest PE_ratio
Cluster 3- has the highest rev_growth and the lowest PE_ratio, asset_turnover
Cluster 4- has the highest PE_ratio
Cluster 5- has the highest net_profit_margin and the lowest Beta 


# Therefore, clusters 1,3 and 5 have the most moderate buying recommendation and the clusters 2 and 4 have hold recommendation


# d)Provide an appropriate name for each cluster using any or all of the variables in the dataset.
Cluster 1 - is the hold cluster
Cluster 2 - is the hold-buy cluster
Cluster 3 - is the buy-sell cluster
Cluster 4 - is the strong buy-sell-hold cluster
Cluster 5 - is the hold-buy cluster



